### 2014-11-08 paper/paths.tex
renamed a file

\section*{Types}

Values in Patina can be described by the following types:
\[
\begin{array}{lccl}
\textrm{Lifetime} & \lt & & \\
\textrm{Type Variable} & X & & \\
\textrm{Qualifier} & q & \bnfdef & \qimm \bnfalt \qmut \\
\textrm{Type} & \gt & \bnfdef & \tyint \bnfalt \own{\gt} \bnfalt \tyref{\lt}{q}{\gt} \bnfalt 
				\subvar{\gt}{i}{n} \bnfalt \subrec{\gt}{i}{n} \bnfalt 
				\tyfix{X}{\gt} \bnfalt X \\
\end{array}
\]

$\own{\gt}$ is a unique pointer to a $\gt$.
$\tyref{\lt}{q}{\gt}$ is a borrowed reference to a $\gt$
providing mutability guarantee $q$ for lifetime $\lt$.
$\subvar{\gt}{i}{n}$ is a variant type.
$\subrec{\gt}{i}{n}$ is a tuple type.
Finally, $\tyfix{X}{\gt}$ is a recursive type.

\section*{L-values}

L-values in Patina are a combination of a variable and a path.
Paths are relative and specify subections of memory reachable from a L-value.

\[
\begin{array}{lccl}
\textrm{Variable} & x & & \\
\textrm{Path} & p & \bnfdef & \base \bnfalt \deref{p} \bnfalt \proj{p}{i} \bnfalt \unroll{\gt}{p} \\
\end{array}
\]

Projection ($\proj{p}{i}$) deconstructs tuples.
Unrolling ($\unroll{\gt}{p}$) deconstructs recursive types.
Dereference ($\deref{p}$) deconstructs pointers and references.
The base path ($\base$) does not deconstruct anything.

The typing judgment for paths does not present any surprises.
We use a partial map for the typing context and
the type substitution operation is the standard capture-avoiding substition.

$$ \gG : \mathrm{Variable} \to \mathrm{Type} $$

\fbox{$\tc{\gG}{\lv{x}{p}}{\gt}$}

\begin{mathpar}
\infer[PT-BASE]{\gG(x)=\gt}{\tc{\gG}{\lv{x}{\base}}{\gt}} \and
\infer[PT-DEOWN]{\tc{\gG}{\lv{x}{p}}{\own{\gt}}}{\tc{\gG}{\lv{x}{\deref{p}}}{\gt}} \and
\infer[PT-DEREF]{\tc{\gG}{\lv{x}{p}}{\tyref{\lt}{q}{\gt}}}{\tc{\gG}{\lv{x}{\deref{p}}}{\gt}} \and
\infer[PT-PROJ]{\tc{\gG}{\lv{x}{p}}{\subrec{\gt}{i}{n}}}{\tc{\gG}{\lv{x}{\proj{p}{i}}}{\gt_i}} \and
\infer[PT-UNROLL]
{\tc{\gG}{\lv{x}{p}}{\tyfix{X}{\gt}}}
{\tc{\gG}{\lv{x}{\unroll{\tyfix{X}{\gt}}{p}}}{\sub{X}{\tyfix{X}{\gt}}{\gt}}}
\end{mathpar}

Unsurprisingly, path typing is unique.
\begin{lem}[Path Type Uniqueness]
  If $\tc{\gG}{\lv{x}{p}}{\gt}$ and $\tc{\gG}{\lv{x}{p}}{\gt'}$ then $\gt = \gt' $.
\end{lem}

\begin{proof}[Path Type Uniqueness]
  We will use induction on the derivation of $\tc{\gG}{\lv{x}{p}}{\gt}$.
  \begin{itemize}
    \item[\textsc{PT-BASE}] Then $p = \base$.

      By inversion, $\gG(x)=\gt$.

      There is only one possible rule (\textsc{PT-BASE}) for 
      deriving $\tc{\gG}{\lv{x}{\base}}{\gt'}$

      Ergo, by inversion, $\gG(x)=\gt'$.

      Thus, $\gt = \gt'$.
    \item[\textsc{PT-DEOWN}] Then $p = \deref{p'}$.

      By inversion, $\tc{\gG}{\lv{x}{p'}}{\own{\gt}}$.

      There are two possible rules for deriving $\tc{\gG}{\lv{x}{\deref{p'}}}{\gt'}$.
      \begin{itemize}
	\item[\textsc{PT-DEOWN}]
	  By inversion, $\tc{\gG}{\lv{x}{p'}}{\own{\gt'}}$.

	  By induction $\own{\gt} =\ \own{\gt'}$.

	  Ergo $\gt = \gt'$.
	\item[\textsc{PT-DEREF}]
	  By inversion, $\tc{\gG}{\lv{x}{p'}}{\tyref{\lt}{q}{\gt'}}$.

	  By induction $\own{\gt} = \tyref{\lt}{q}{\gt'}$, a contradiction.

	  Thus, this case is impossible.
      \end{itemize}
    \item[\textsc{PT-DEREF}] Then $p = \deref{p'}$.

      Similar to the \textsc{PT-DEOWN} case, 
      but swapping the roles of \textsc{PT-DEOWN} and \textsc{PT-DEREF}.
    \item[\textsc{PT-PROJ}] Then $p = \proj{p'}{i}$ and $\gt = \gt_i$.

      By inversion, $\tc{\gG}{\lv{x}{p'}}{\subrec{\gt}{i}{n}}$.

      There is only one rule (\textsc{PT-PROJ}) for deriving $\tc{\gG}{\lv{x}{\proj{p'}{i}}}{\gt'}$.

      Thus, by inversion, $\gt' = \gt'_i$ and $\tc{\gG}{\lv{x}{p'}}{\subrec{\gt'}{i}{n'}}$.

      By induction, $\subrec{\gt}{i}{n} = \subrec{\gt'}{i}{n'}$.

      Ergo $n = n'$ and $\gt_i = \gt'_i$.
    \item[\textsc{PT-UNROLL}] 
      Then $p = \unroll{\tyfix{X}{\gt_0}}{p'}$ and $\gt = \sub{X}{\tyfix{X}{\gt_0}}{\gt_0}$.

      There is only one rule (\textsc{PT-UNROLL}) for deriving
      $\tc{\gG}{\lv{x}{\unroll{\tyfix{X}{\gt_0}}{p'}}}{\gt'}$.

      Thus, by inversion, $\gt' = \sub{X}{\tyfix{X}{\gt_0}}{\gt_0}$.

      Ergo $\gt = \sub{X}{\tyfix{X}{\gt_0}}{\gt_0} = \gt'$.
  \end{itemize}
\end{proof}

\section*{Runtime Memory}

\subsection*{Representation and Addressing}

To accurately model Rust's memory usage, Patina restricts
the contents of a memory cell to void data, an integer, or a pointer to another cell.
Tuples and recursive types have no physical memory presence (beyond contiguity for tuples).
The memory representation of a variant is a pair of a cell for the discriminant
and whatever memory is necessary to store the payload.

We could model this by a map from addresses to memory cell values,
but two issues make this inconvenient: the need for address arithmetic
and non-unique typing. However, we can add a little extra structure to
our model and eliminate these two issues. We wrap the cells inside
layouts describing the type structure overlaying memory.

We also separate plain pointer cells into owned cells and reference cells.
This is mostly useful for providing just enough type information about the pointer
for memory operation purposes. We will explain how we address memory in a moment,
so we use a placeholder for now.

\[
\begin{array}{lccl}
\textrm{Integer} & z & & \\
\textrm{Address} & \rho & & \textrm{a placeholder} \\
\textrm{Cell} & c & \bnfdef & \void \bnfalt z \bnfalt \own{\rho} \bnfalt \Ref{q~\rho} \\
\textrm{Layout} & l & \bnfdef & c \bnfalt \varval{c}{l} \bnfalt 
			       \subrec{l}{i}{n} \bnfalt \roll{\gt}{l} \\
\end{array}
\]

Due to the extra structure from layouts, addressing memory now requires
more than simple labels. Instead we use a runtime analogue of variables and paths.
Allocations are chunks of memory allocated and freed atomically.
They correspond to either variables or heap allocations.
Routes provide relative specification into layouts similar to paths with two key differences.

\[
\begin{array}{lccl}
\textrm{Allocation} & \ga & & \\
\textrm{Route} & r & \bnfdef & \base \bnfalt \proj{r}{i} \bnfalt \pay{r} \bnfalt \unroll{\gt}{r} \\
\end{array}
\]

The $\base$ route, projection, and unrolling are effectively identical to their path equivalents.
The $\pay{r}$ route refers to the payload $l$ in a variant $\varval{c}{l}$.
This is primarily used for match-by-reference.
Unlike paths, there is no dereference route.
This forces any pointer following into path evaluation (into routes)
rather than doing so while reading memory at a route.
The address placeholder from before is simply a pair of an allocation label and a route.
That is:

\[
\begin{array}{lccl}
\textrm{Integer} & z & & \\
\textrm{Allocation} & \ga & & \\
\textrm{Route} & r & \bnfdef & \base \bnfalt \proj{r}{i} \bnfalt \pay{r} \bnfalt \unroll{\gt}{r} \\
\textrm{Cell} & c & \bnfdef & \void \bnfalt z \bnfalt 
			      \own{\addr{\ga}{r}} \bnfalt \refval{q}{\addr{\ga}{r}} \\
\textrm{Layout} & l & \bnfdef & c \bnfalt \varval{c}{l} \bnfalt 
			       \subrec{l}{i}{n} \bnfalt \roll{\gt}{l} \\
\end{array}
\]

\subsection*{Reading}

Reading a layout from a route in memory is a straightforward operation.
We model the heap as a partial map from allocations to layouts.
Routes and layouts interact as you would expect.

$$ H : \mathrm{Allocation} \to \mathrm{Layout} $$

\fbox{$\Read{H}{\addr{\ga}{r}}{l}$}

\begin{mathpar}
\infer[RD-BASE]{H(\ga)=l}{\Read{H}{\addr{\ga}{\base}}{l}} \and
\infer[RD-PROJ]
  {\Read{H}{\addr{\ga}{r}}{\subrec{l}{i}{n}}}
  {\Read{H}{\addr{\ga}{\proj{r}{i}}}{l_i}} \and
\infer[RD-PAY]{\Read{H}{\addr{\ga}{r}}{\varval{c}{l}}}{\Read{H}{\addr{\ga}{\pay{r}}}{l}} \and
\infer[RD-UNROLL]{\Read{H}{\addr{\ga}{r}}{\roll{\gt'}{l}}}{\Read{H}{\addr{\ga}{\unroll{\gt}{r}}}{l}}
\end{mathpar}

As one would expect of a read operation, it's result is unique.

\begin{lem}[Read Uniqueness]
  If $\Read{H}{\addr{\ga}{r}}{l}$ and $\Read{H}{\addr{\ga}{r}}{l'}$, then $l = l'$.
\end{lem}

\begin{proof}[Read Uniqueness]
  We will induct on the derivation of $\Read{H}{\addr{\ga}{r}}{l}$.
  \begin{itemize}
    \item[\textsc{RD-BASE}] Then $r = \base$.

      By inversion, $H(\ga)=l$.

      There is only one rule (\textsc{RD-BASE}) for deriving $\Read{H}{\addr{\ga}{\base}}{l'}$.

      Thus, by inversion, $H(\ga)=l'$.

      Ergo, $l = l'$.
    \item[\textsc{RD-PROJ}] Then $r = \proj{r'}{i}$ and $l = l_i$.

      By inversion, $\Read{H}{\addr{\ga}{r'}}{\subrec{l}{i}{n}}$.

      There is only one rule (\textsc{RD-PROJ}) for
      deriving $\Read{H}{\addr{\ga}{\proj{r'}{i}}}{l'}$.

      Thus, by inversion, $\Read{H}{\addr{\ga}{r'}}{\subrec{l'}{i}{n'}}$.

      By induction, $\subrec{l}{i}{n} = \subrec{l'}{i}{n'}$.

      Ergo, $n = n'$ and $l_i = l'_i$.
    \item[\textsc{RD-PAY}] Then $r = \pay{r'}$.

      By inversion, $\Read{H}{\addr{\ga}{r'}}{\varval{c}{l}}$.

      There is only one rule (\textsc{RD-PAY}) for deriving $\Read{H}{\addr{\ga}{\pay{r'}}}{l'}$.

      Thus, by inversion, $\Read{H}{\addr{\ga}{r'}}{\varval{c'}{l'}}$.

      By induction, $\varval{c}{l} = \varval{c'}{l'}$.

      Ergo, $c = c'$ and $l = l'$.
    \item[\textsc{RD-UNROLL}] Then $r = \unroll{\gt}{r'}$.

      By inversion, $\Read{H}{\addr{\ga}{r'}}{\roll{\gt'}{l}}$.

      There is only one rule (\textsc{RD-UNROLL}) for
      deriving $\Read{H}{\addr{\ga}{\unroll{\gt}{r'}}}{l'}$.

      Thus, by inversion, $\Read{H}{\addr{\ga}{r'}}{\roll{\gt''}{l'}}$.

      By induction, $\roll{\gt'}{l} = \roll{\gt''}{l'}$.

      Ergo, $\gt' = \gt''$ and $l = l'$.
  \end{itemize}
\end{proof}

\section*{Path Evaluation}

Now that we can read memory, we can define path evaluation.
The only actual work of path evaluation is following the dereferences of pointers
to produce a route. In order to connect variables to runtime allocations, we use
a partial map tracking the allocation labels of variables.

$$ V : \mathrm{Variable} \to \mathrm{Allocation} $$

\fbox{$\ev{V;H}{\lv{x}{p}}{\addr{\ga}{r}}$}

\begin{mathpar}
\infer[PE-BASE]{V(x)=\ga}{\ev{V;H}{\lv{x}{\base}}{\addr{\ga}{\base}}} \and
\infer[PE-DEOWN]
  {\ev{V;H}{\lv{x}{p}}{\addr{\ga}{r}} \\ \Read{H}{\addr{\ga}{r}}{\own{\addr{\ga'}{r'}}}}
  {\ev{V;H}{\lv{x}{\deref{p}}}{\addr{\ga'}{r'}}} \and
\infer[PE-DEREF]
  {\ev{V;H}{\lv{x}{p}}{\addr{\ga}{r}} \\ \Read{H}{\addr{\ga}{r}}{\refval{q}{\addr{\ga'}{r'}}}}
  {\ev{V;H}{\lv{x}{\deref{p}}}{\addr{\ga'}{r'}}} \and
\infer[PE-PROJ]
  {\ev{V;H}{\lv{x}{p}}{\addr{\ga}{r}}}
  {\ev{V;H}{\lv{x}{\proj{p}{i}}}{\addr{\ga}{\proj{r}{i}}}} \and
\infer[PE-UNROLL]
  {\ev{V;H}{\lv{x}{p}}{\addr{\ga}{r}}}
  {\ev{V;H}{\lv{x}{\unroll{\gt}{p}}}{\addr{\ga}{\unroll{\gt}{r}}}}
\end{mathpar}

With path typing and evaluation now defined, we would like to prove progress
and preservation for them. Preservation will require we define typing for
runtime data (routes, cells, and layouts), which involves a slight bit of trickery
around variant discriminants and the payload route.

Progress will require a way to guarantee that the necessary \textsc{read}s 
can be completed, i.e. any pointers we need to dereference are in fact initialized
(a property we call \emph{shallow initialization}).
We will accomplish this by using a kind of shadow heap, which will simply track
initialization rather than values.
This will require an analogue of \textsc{read} that extracts the initialization
data for a path, a proof that this operation preserves types 
(which itselfs requires typing for this shadow heap),
a coherence check to ensure the shadow heap models the runtime heap,
and a proof that a shallowly initialized path evaluates to a readable route.