### 2014-11-12 paper/intro.tex
through path type preservation

%\section*{Introduction}
%Rust claims to ensure memory safety without a garbage collector.
%We would like to verify this claim.

%Rust ensures safety by combining affine types with checked borrowed references.
%Affine types ensure that every piece of memory has a unqiue owner responsible for deallocation.
%Borrowed references allow limited aliasing that avoids 
%aliased, mutable memory and dangling references
%Rust's usage of affine types for memory management is standard except for one addition.
%Rust allows consumed affine cells that are still in scope to be re-initialized and used again.
%Rust's borrowed references utilize a region system very similar to Cyclone's.
%The primary difference is in the details of the borrow checker.
%Our main goal is to prove that these differences do not introduce unsoundness.

%\section*{Patina}
%We work with a small core model of Rust called Patina, which focuses
%on the memory management aspects of Rust.
%Unlike Rust, Patina uses explicit heap deallocation and does not support closures.
%Patina has three layers: L-values, R-values, and statements.
%We will discuss them in order.

\section*{Types}

Values in Patina can be described by the following types:
\[
\begin{array}{lccl}
\textrm{Lifetime} & \lt & & \\
\textrm{Type Variable} & X & & \\
\textrm{Qualifier} & q & \bnfdef & \qimm \bnfalt \qmut \\
\textrm{Type} & \gt & \bnfdef & \tyint \bnfalt \own{\gt} \bnfalt \tyref{\lt}{q}{\gt} \bnfalt 
				\subvar{\gt}{i}{n} \bnfalt \subrec{\gt}{i}{n} \bnfalt 
				\tyfix{X}{\gt} \bnfalt X \\
\end{array}
\]

$\own{\gt}$ is a unique pointer to a $\gt$.
$\tyref{\lt}{q}{\gt}$ is a borrowed reference to a $\gt$
providing mutability guarantee $q$ for lifetime $\lt$.
$\subvar{\gt}{i}{n}$ is a variant type.
$\subrec{\gt}{i}{n}$ is a tuple type.
Finally, $\tyfix{X}{\gt}$ is a recursive type.