### 2015-02-14 paper/stmts.tex
statements file

\section*{Statements}
\subsection*{Statement Typing}
\subsection*{Statement Evaluation}
\subsection*{Statement Preservation}
\subsection*{Statement Progress}
\section*{Patina Soundness}