### 2015-02-16 paper/stmts.tex
statement preservation and progress

\section*{Statements}
We have finally reached the outermost shell of Patina!
The statements layer is primarily about heap manipulation and chaining expressions.
It mostly just relies on judgments we have already defined.
$\nop$ and $\seq{s}{s}$ are simply for chaining together statements.
$\set{\lv{x}{p}}{e}$ is for initializing a path with the result of an expression.
$\new{\lv{x}{p}}{\lv{x}{p}}$ is similar, but allocates the result on the heap 
and initializes the path with a new unique pointer. 
Treating this as a statement rather than an expression
simplifies the expression preservation and progress lemmas.
Note that both versions are for \emph{initialization} and not \emph{assignment}.
We only allow initialization of droppable paths, which we guarantee will not orphan memory.
$\swap{\lv{x}{p}}{\lv{x}{p}}$ will perform a shallow swap of two writable paths.
This is the only way to actually mutate through a mutable borrow since we do not allow
interior of a borrowed reference to be uninitialized even temporarily.
$\free{\lv{x}{p}}$ performs a shallow free and is the inverse of \textsf{new}.
$\push{\lt}{x}{\gt}{s}$ will create a new stack variable.
$\pop{x}{s}$ will pop that stack variable once the statement that requires it finishes.
It is for evaluation purposes and should not appear in user code.

\[
\begin{array}{lccl}
\textrm{Statements} & s & \bnfdef & \nop \bnfalt \seq{s}{s} \bnfalt \set{\lv{x}{p}}{e}
			  \bnfalt \new{\lv{x}{p}}{\lv{x}{p}} \bnfalt \swap{\lv{x}{p}}{\lv{x}{p}}
			  \bnfalt \free{\lv{x}{p}} \bnfalt \push{\lt}{x}{\gt}{s}
			  \bnfalt \pop{x}{s}
\end{array}
\]

\subsection*{Droppable Shadows}
Earlier we defined $\textsc{droppable}(l)$, which specified layouts that were safe to deallocate.
Now we define an equivalent notion for shadows, which we will use to ensure only droppable
layouts are deallocated. We also require a droppable shadow to be unborrowed, which will
prevent us from creating dangling borrowed references.
\newline

\fbox{$\gs\ \textsc{droppable}$}

\begin{mathpar}
  \infer[DS-UNINIT]{ }{\emptyset : \uninit\ \textsc{droppable}} \and
  \infer[DS-INT]{ }{\emptyset : \tyint\ \textsc{droppable}} \and
  \infer[DS-REF]
    {\cnl{\gs}}
    {\emptyset : \refval{q}{\gs}\ \textsc{droppable}}
\end{mathpar}

\subsection*{Subtyping}
The only subtyping in Patina is lifetime subtyping.
Borrowed references are contravariant in their lifetimes.
This means that a reference for a longer lifetime is a 
subtype of a reference for a shorter lifetime, which is the behavior we want.
The reference of a longer lifetime is usable anywhere the reference of a shorter lifetime is.
Since mutable references can both read and write, they are invariant in their type argument.
Immutable references can only read, so they are covariant in their type argument.
\newline

\fbox{$\ltr\vdash\gt_0<:\gt_1$}

\begin{mathpar}
  \infer[SUB-INT]{ }{\ltr\vdash\tyint<:\tyint} \and
  \infer[SUB-OWN]{\ltr\vdash\gt_1<:\gt_2}{\ltr\vdash\ \own{\gt_1}<:\ \own{\gt_2}} \and
  \infer[SUB-REFIMM]
    {(\lt_2,\lt_1)\in\ltr \\ \ltr\vdash\gt_1<:\gt_2}
    {\ltr\vdash\tyref{\lt_1}{\qimm}{\gt_1}<:\tyref{\lt_2}{\qimm}{\gt_2}} \and
  \infer[SUB-REFMUT]
    {(\lt_2,\lt_1)\in\ltr}
    {\ltr\vdash\tyref{\lt_1}{\qmut}{\gt}<:\tyref{\lt_2}{\qmut}{\gt}}
\end{mathpar}

\subsection*{Statement Typing}
\fbox{$\tc{\gG;\ltr;L;\gU}{s}{\gU'}$}
\newline

The $\nop$ statement does nothing so it is easy to check.

\begin{mathpar}
  \infer[ST-SKIP]{ }{\tc{\gG;\ltr;L;\gU}{\nop}{\gU}}
\end{mathpar}

Sequencing two statement simply sequences typechecking by threadings the shadow heap.
Note that since $\gG$ is not changing, the statements can only change the initialization
and loan information of the shadow heap.

\begin{mathpar}
  \infer[ST-SEQ]
    {\tc{\gG;\ltr;L;\gU_0}{s_1}{\gU_1} \\ \tc{\gG;\ltr;L;\gU_1}{s_2}{\gU_2}}
    {\tc{\gG;\ltr;L;\gU_0}{\seq{s_1}{s_2}}{\gU_2}}
\end{mathpar}

Swapping requires only that the two L-values have the same type and that they both
be writable. Since the writability guarantees no part is loaned, the shadow heap is not changed.

\begin{mathpar}
  \infer[ST-SWAP]
    {\tc{\gG}{\lv{x_1}{p_1}}{\gt} \\ \cw{p_1}{\gU(x_1)} \\\\
     \tc{\gG}{\lv{x_2}{p_2}}{\gt} \\ \cw{p_2}{\gU(x_2)}}
    {\tc{\gG;\ltr;L;\gU}{\swap{\lv{x_1}{p_1}}{\lv{x_2}{p_2}}}{\gU}}
\end{mathpar}

Initializing with a value requires that the path be droppable so that nothing will be orphaned.
It also requires the expression type to be a subtype of the path type.
As discussed above, this refers to lifetimes: the subtype lifetime is greater than or
equal to the lifetime of the supertype. This requirement guarantees that the value
stored at the path will be valid for at least as long as the path is valid.
Finally, we initialize the path in the output shadow heap.

\begin{mathpar}
  \infer[ST-SET]
    {\tc{\gG}{\lv{x}{p}}{\gt_p} \\
     \tc{\gG;\ltr;L;\gU_0}{e}{\gt_e;\gU_1} \\
     \ltr\vdash\gt_e<:\gt_p \\\\
     \shallow{\gU_1}{\lv{x}{p}}{\gs} \\
     \gs\ \textsc{droppable} \\
     \textsc{init}(\gU_1,\lv{x}{p},\gU_2)
    }
    {\tc{\gG;\ltr;L;\gU_0}{\set{\lv{x}{p}}{e}}{\gU_2}}
\end{mathpar}

Initializing with a pointer is similar to initializing with a value.
Except instead of expression typing, we must check that we can use the path 
on the right hand side.

\begin{mathpar}
  \infer[ST-NEW]
    {\tc{\gG}{\lv{x_l}{p_l}}{\own{\gt_l}} \\
     \uselv{\gG}{\gU_0}{\lv{x_r}{p_r}}{\own{\gt_r}}{\gU_1} \\
     \ltr\vdash\ \own{\gt_r}<:\ \own{\gt_l} \\\\
     \shallow{\gU_1}{\lv{x_l}{p_l}}{\gs} \\
     \gs\ \textsc{droppable} \\
     \textsc{init}(\gU_1,\lv{x_l}{p_l},\gU_2)
    }
    {\tc{\gG;\ltr;L;\gU_0}{\new{\lv{x_l}{p_l}}{\lv{x_r}{p_r}}}{\gU_2}}
\end{mathpar}

Freeing the memory of a unique pointer requires checking that that memory is droppable.
We then mark it used in the output shadow heap, which since this a unique pointer
will replace the shadow with $\uninit$.

\begin{mathpar}
  \infer[ST-FREE]
  {\tc{\gG}{\lv{x}{p}}{\own{\gt}} \\
   \shallow{\gU}{\lv{x}{p}}{\own{\gs}} \\
   \gs\ \textsc{droppable} \\
   \textsc{use}(\gU,\lv{x}{p},\gU')
  }
  {\tc{\gG;\ltr;L;\gU}{\free{\lv{x}{p}}}{\gU'}}
\end{mathpar}

Pushing a new stack variable involves checking that the body is well typed
under the appropriately extended contexts. Newly pushed variables start uninitialized.
By the end of the body, the variable must be droppable so that popping it will not
orphan memory. When checking the body, we extend the lifetime relation with the new lifetime
by adding the new lifetime as the smallest lifetime in the relation (it is related to everything
else). Afterwards, we must filter everything involving this lifetime from the shadow heap.
This should only consist of removing loans from banks. We also "deallocate" the stack variable
from the shadow heap by restricting the domain to the domain of the input.

\begin{mathpar}
  \infer[ST-PUSH]
  {\textsc{new-smallest-lifetime}(\ltr,\lt,\ltr') \\\\
   \tc{\gG,x\mapsto\gt;\ltr';L,x\mapsto\lt;\gU_0,x\mapsto\uninit}{s}{\gU_1} \\\\
   \gU_1(x)\ \textsc{droppable} \\
   \textsc{filter}(\gU_1,\lt,\gU_2)
  }
  {\tc{\gG;\ltr;L;\gU_0}{\push{\lt}{x}{\gt}{s}}{\gU_2|_{\dom \gU_0}}}
\end{mathpar}

Popping a stack variable is similar, but we do not need to worry about extending contexts.
Instead, we must make sure to remove the stack variable when we restrict the domain
of the output shadow heap.

\begin{mathpar}
  \infer[ST-POP]
  {\tc{\gG;\ltr;L;\gU_0}{s}{\gU_1} \\\\
   \gU_1(x)\ \textsc{droppable} \\
   \textsc{filter}(\gU_1,\lt,\gU_2)
  }
  {\tc{\gG;\ltr;L;\gU_0}{\pop{x}{s}}{\gU_2|_{\dom \gU_0 - \{x\}}}}
\end{mathpar}

One interesting thing to note here is that the domain of the input and output shadow heaps
is always the same except for the \textsf{pop} case, where it decreases by one.
Since \textsf{pop} does not appear in user code, this means that if a program starts
with no memory allocated and that program type checks, then the program frees all the
memory it allocates. Assuming of course that evaluation preserves the shadow heap.

\subsection*{Statement Evaluation}
Statement evaluation goes as you would expect.
We use a new helper judgment $\textsc{fresh}(\ga)$ that generates a unused allocation.
As before, we utilize domain restriction to "deallocate".
Other than that, we simply use path and expression evaluation combined
with \textsc{read} and \textsc{write}.
\newline

\fbox{$\sev{V}{H}{s}{V'}{H'}{s'}$}

\begin{mathpar}
  \infer[SE-SKIPSEQ]{ }{\sev{V}{H}{\seq{\nop}{s}}{V}{H}{s}} \and
  \infer[SE-STEPSEQ]
    {\sev{V}{H}{s_1}{V'}{H'}{s'_1}}
    {\sev{V}{H}{\seq{s_1}{s_2}}{V'}{H'}{\seq{s'_1}{s_2}}} \and
  \infer[SE-SET]
    {\ev{V;H_0}{e}{l;H_1} \\
     \ev{V;H_1}{\lv{x}{p}}{\ga} \\
     \textsc{write}(H_1,\ga,l,H_2)
    }
    {\sev{V}{H_0}{\set{\lv{x}{p}}{e}}{V}{H_2}{\nop}} \and
  \infer[SE-NEW]
    {\ev{V;H_0}{\lv{x_r}{p_r}}{\ga_r} \\
     \Read{H_0}{\ga_r}{l_r} \\
     \textsc{write}(H_0,\ga_r,\textsc{use}(l_r),H_1) \\\\
     \ev{V;H_1}{\lv{x_l}{p_l}}{\ga_l} \\
     \textsc{fresh}(\ga_n) \\
     \textsc{write}(H_1,\ga_l,\own{\ga_n},H_2) \\
    }
    {\sev{V}{H_0}{\new{\lv{x_l}{p_l}}{\lv{x_r}{p_r}}}{V}{H_2, \ga_n\mapsto l_r}{\nop}} \and
  \infer[SE-SWAP]
    {\ev{V;H_0}{\lv{x_l}{p_l}}{\ga_l} \\
     \ev{V;H_0}{\lv{x_r}{p_r}}{\ga_r} \\\\
     \Read{H_0}{\ga_l}{l_l} \\
     \Read{H_0}{\ga_r}{l_r} \\\\
     \textsc{write}(H_0,\ga_l,l_r,H_1) \\
     \textsc{write}(H_1,\ga_r,l_l,H_2)
    }
    {\sev{V}{H_0}{\swap{\lv{x_l}{p_l}}{\lv{x_r}{p_r}}}{V}{H_2}{\nop}} \and
  \infer[SE-FREE]
    {\ev{V;H_0}{\lv{x}{p}}{\ga} \\
     \Read{H_0}{\ga}{\own{\ga'}} \\
     \textsc{write}(H_0,\ga,\textsc{use}(\own{\ga'}),H_1)
    }
    {\sev{V}{H_0}{\free{\lv{x}{p}}}{V}{H_1|_{\dom H_1 - \{\ga'\}}}{\nop}} \and
  \infer[SE-PUSH]
    {\textsc{fresh}(\ga)}
    {\sev{V}{H}{\push{\lt}{x}{\gt}{s}}{V,x\mapsto\ga}{H,\ga\mapsto\void}{\pop{x}{s}}} \and
  \infer[SE-POPSTEP]
    {\sev{V}{H}{s}{V'}{H'}{s'}}
    {\sev{V}{H}{\pop{x}{s}}{V'}{H'}{\pop{x}{s'}}} \and
  \infer[SE-POPSKIP]
    { }
    {\sev{V}{H}{\pop{x}{\nop}}{V|_{\dom V - \{x\}}}{H|_{\dom H - \{V(x)\}}}{\nop}}
\end{mathpar}

\subsection*{Statement Preservation}
Statement preservation is the assertion that given all our well-formed contexts,
a well-typed statement, and the fact that that statement evaluates to another statement,
then we can also type the result statement and both statements produce the same output
shadow heap, i.e.\ they effectively end in the same way.
Put another way, this lemma asserts that memory safe programs step to memory safe programs.

It is difficult to specify how the new versions of contexts relate to the old versions.
They may be larger or smaller depending upon the statement in question.
We do not specify this relationship currently.
Instead we just rely on the new variable map, heap and statement to enforce
the relationship.

\begin{conj}[Statement Preservation]
  If $\tc{}{H}{\gS}$, $\gG;V\vdash\gS$, $\vdash\ltr$, $\tb{\ltr}{\gU}{\gG}{L}$,
  $\tc{V}{H}{\gU}$, $\tc{\gG;\ltr;L;\gU}{s}{\gU_0}$, and $\sev{V}{H}{s}{V'}{H'}{s'}$
  then there exist $\gS'$, $\gG'$, $\ltr'$, $L'$, and $\gU'$ such that
  $\tc{}{H'}{\gS'}$, $\gG';V'\vdash\gS'$, $\vdash\ltr'$, $\tb{\ltr'}{\gU'}{\gG'}{L'}$,
  $\tc{V'}{H'}{\gU'}$, and $\tc{\gG';\ltr';L';\gU'}{s'}{\gU_0}$.

\end{conj}

\begin{proof}[Justification]
  The skip-seq case is trivial.
  The step-seq case follows from induction.
  In the value initialization (set) case, only the shadow heap and the heap are changing.
  From Expression Preservation, we know that the heap and shadow heap are still aligned
  after evaluating the expression. The \textsc{write} will initialize the path, which
  is reflected in the shadow heap.
  The pointer initialization (new) case is similar.
  Since the components of a swap must have the same shadow, the modified heap will
  still be well-formed.
  We required that the memory deallocated by \textsf{free} was droppable, which
  should mean that the heap is still well-formed once we $\void$ the pointer.
  Push steps immediately to pop and only extends contexts, which should still be well-formed.
  The pop-step case follows from induction.
  In the pop-skip case, similar to the \textsf{free} case, we know that the variable
  is droppable, which should mean the resulting contexts are well-formed.
\end{proof}

\subsection*{Statement Progress}
Statement progress is essentially the assertion that a well-typed statement will
never attempt to read uninitialized memory, free borrowed memory, and so on.
Essentially, that well-typed programs have sufficient memory safety to take a step.
The restriction to statements that end with an empty shadow heap allows us
to avoid the cases with orphan memory.

\begin{conj}[Statement Progress]
  If $\tc{}{H}{\gS}$, $\gG;V\vdash\gS$, $\vdash\ltr$, $\tb{\ltr}{\gU}{\gG}{L}$,
  $\tc{V}{H}{\gU}$, $\tc{\gG;\ltr;L;\gU}{s}{\emptyset}$
  then there exists $V'$, $H'$, and $s'$ such that $\sev{V}{H}{s}{V'}{H'}{s'}$
  or $s = \nop$.
\end{conj}

\begin{proof}[Justification]
  The skip case is immediate.
  The sequence case is either immediate or follows from induction.
  The swap case and the new case follow from path progress and read safety.
  The set case is similar, but also requires expression progress.
  The push case is immediate.
  The pop case is either immediate or follows from induction.
\end{proof}

\section*{Patina Soundness}
If our conjectures hold, then we have established soundness for Patina.
This means that well-typed statement that ends with an empty shadow heap
will not leak memory, not access uninitialized memory, not create aliased, mutable memory,
and so on. In effect, well-typed programs are statically guaranteed to be memory safe.