### 2015-02-26 paper/futurework.tex

\section*{Future Work}
The first extension to the existing work should be to prove our conjectures.
This is necessary to put us on a firm footing for other future work, and
to ensure we have actually included everything necessary in our model.

The expression and statement layers should be extended to match the path layer.
Product types are straightforward to add to the borrow checker and will like not
pose much trouble. Sum types may be more difficult. In particular, matching by
reference allows creation of references to the payload of a sum type.
The borrow checker must ensure that the sum is not mutated while such a reference
exists. Extending the model with recursive types will require the addition of
top level (non-closure) recursive functions. Without the capability those functions
give us it is impossible to free recursive values. Those functions will necessitate
other additions, e.g. lifetime parameters. Type parameters likely do not have
significant soundness implications by themselves, but may in their interaction with
later features.

Until recently, Rust's closures have been in flux. As of this writing, Rust supports
both stack and heap closures and also supports different manners of access to the environment
(owning, immutable reference, or mutable reference). In the past, closures have been a
common source of borrow checker bugs, which likely means that proving soundness will be
difficult. However, it also means that it is one of the most essential features to add to Patina.
It would likely be profitable to look to Cyclone \cite{cycregions}, which took great
care to ensure references could not escape via closures or existential types.

Traits, which are Rust's version of typeclasses or interfaces, may have similar issues.
Traits can be used to create what is effectively an object or an existential type.
Since objects and closures are so similar, it is likely that proving trait objects sound
will face similar difficulties. Rust's closures are actually implemented as syntactic sugar
for certain trait implementations in the current version. 
It is unclear whether this means formalization of closures will be easier
or whether formalization of traits will be harder.

Recently, Rust has added some more advanced type system features.
In particular, associated types and higher rank lifetimes.
Associated types are effectively higher kinded types.
Neither of these features will likely impact soundness, but they should still be
carefully modeled.

There are many potential features that we could explore using our model.
Several additional versions of borrowed references has been purposed,
e.g. write-only references or references that allow moving out as long as something
is moved back in before the reference is released. Adding this to Patina would likely
result in a more general borrow checking framework.

One oft-requested feature is the ability to use static values in type parameters.
This effectively gives Rust a weak form of dependent typing.
The consequences for soundness in a affine dependently typed language with regions
are likely significant, and the difficulty in proving them even more so.

Our model made several changes to how systems worked in the Rust compiler.
It would be interesting to compare how the model's version compares with the original
in terms of performance, ease of understanding, and mantainability.